Checking Big Suffix and LCP Arrays by Probabilistic Methods

### 2017-03-20 draft-old/draft.tex
2017/03/20

For full-text indexing of massive data, the suffix and LCP (longest common prefix) arrays have been recognized as fundamental data structures, and there are at least two needs in practice for checking their correctness, i.e. program debugging and verifying the arrays constructed by probabilistic algorithms. In this paper, we propose two probabilistic methods to check the suffix and LCP arrays of constant or integer alphabets in external memory by using a Karp-Rabin fingerprinting technique, where the checking result is wrong only with a negligible error probability. The first method checks the lexicographical order and the LCP-value of two suffixes by computing and comparing the fingerprints of their LCPs. This method is rather general in terms of that it can verify any full or sparse suffix/LCP array of any order. The second method is more space efficient, it first employs the fingerprinting technique to verify a subset of the given suffix and LCP arrays, from which then a copy of the suffix and LCP arrays is produced by using the induced sorting method and compared with the given arrays for verification. In particular, for constant alphabets, the copy of the induced suffix and LCP arrays can be removed to make a further improvement on the space efficiency.

@@

Our contribution comprises two methods to probabilistically verify any given suffix and LCP arrays. In principle, Method A checks the lexical order and the LCP-value of two neighboring suffixes in the given SA by literally comparing the characters of their LCPs. For reducing the time complexity of a comparison between two sequences of characters, we use a Karp-Rabin fingerprinting technique to convert each sequence into a single integer, called fingerprint, and compare the fingerprints instead to check the equality of two sequences. The algorithm for Method A involves multiple scans and sorts on sets of $\mathcal{O}(n)$ fixed-size items. Its implementation in external memory suffers from a space bottleneck due to the large disk volume taken by each sort. 
To overcome this drawback, Method B first employs the fingerprinting technique to check a subset selected from the given suffix and LCP arrays, then it utilizes the IS method to induce the final suffix and LCP arrays from the verified subset and literally compares them with the input arrays to ensure the correctness of the latter. Our experimental results indicate that the programs for Algorithm~\ref{alg:2} designed by Method B only take around half as much disk space as the program for Algorithm~\ref{alg:1} designed by Method A.

The remainder of this paper is organized as follows. Sections~\ref{sec:method1} and~\ref{sec:method2} describe the two methods and their algorithmic designs. Section~\ref{sec:experiment} conducts an experimental study for performance evaluation of our programs for these algorithms. Section~\ref{sec:conclusion} gives some concluding remarks.

@@

{\bf Assume that the alphabet $\Sigma$ is of a constant size, we can check $sa$ and $lcp$ without producing a copy of the final suffix and LCP arrays in Algorithm~\ref{alg:2}.} The idea is to compare the induced suffix/LCP items with their corresponding items in $sa/lcp$ during the inducing process. Specifically, when a suffix/LCP item $v_1$ is induced into a bucket, we check if it is equal to the corresponding item $v_2$ in $sa/lcp$. If $v_1 = v_2$, then $v_2$ is correct and we further use this value to induce the remaining suffix/LCP items. We describe more details of the adapted inducing process as below, where $lp_1/lp_2$ and $sp_1/sp_2$ point to the next items to be visited in the L-type and S-type sub-buckets, respectively.

@@

We select these solutions because they are the current fastest available to us. A runtime breakdown of these solutions on the prefixes of "enwiki" is given in Table~\ref{tbl:3}, from which we see that Solution 2 is about two times faster than Solution 1 and twice as fast as ProgA. This phenomenon is mainly due to the high speed of pSAscan in the experiment. However, it is worthy of pointing out that both pSAscan and Sparse-$\phi$ are of the time and I/O complexities proportional to $n^2/M$. This is much higher than eSAIS and our checking algorithms when $n$ increases, and thus poses a strict limitation to the scalability of Solution 2. As reported in~\cite{Karkkainen2015}, when $n$ is considerably greater than $M$, eSAIS is much more time and I/O efficient than pSAscan. In this experiment, pSAscan builds the SA for "enwiki\_8g" in time double as that for "enwiki\_1g". {\bf For big $n$, it is more reasonable to compare the results of our programs with that of Solution 1.}

@@

It should be noticed that our programs are coded for experimental study only. From engineering aspects, there is still a big margin for better implementation. For example, several algorithms for induced sorting a suffix and/or LCP arrays were proposed these years \cite{Nong15,Nong14,Bingmann12}, with different methods for solving the key problem of retrieving the preceding character of a sorted suffix in the inducing process. A recent work \cite{Karkkainen2017} for engineering these IS methods with some implementation optimizing techniques achieves a significant improvement over the previous results. As reported, the peak disk use is $7n$ for 32-bit integers. Because the inducing process is the performance bottleneck for Method B, it is reasonable to expect that a better engineering implementation of the method will yield a remarkable performance improvement. {\bf An optimized engineering of our methods is out of the scope of this paper and will be addressed elsewhere.}

@@

The work presented in~\cite{Fischer11} describes how to compute the LCP array during S2-S3 of the induction phase. Specifically, for any two suffixes placed at the neighboring positions in $sa$, their LCP-value can be computed according to one of the following two cases in respect to whether or not they are inserted into the same sub-bucket: if yes, then the LCP-value is one greater than that of the two suffixes from which inducing them; otherwise, the LCP-value is zero. The key operation herein is to compute the LCP-value of the inducing suffixes at the same time when inserting the induced suffixes into $sa$. For example, in lines 10-21 of Fig.~\ref{fig:example2}, we induce ${\sf suf}(12)$ and ${\sf suf}(6)$ into the neighboring positions of ${\sf sa\_bkt_L(2)}$, respectively. If we keep recording the minimal over $lcp(0, 5]$, then we can obtain the LCP-value of the inducing suffixes ${\sf suf}(13)$ and ${\sf suf}(5)$ immediately after putting ${\sf suf}(6)$ into $sa$. This problem is modeled as a range minimum query in~\cite{Fischer11} and can be answered within amortized $\mathcal{O}(1)$ time.

@@

We are intended for ....

put emphasis on

with the same ease as when

The biggest difference between two is that ...

Toward the end, ...

they are designed to be both easy to interpret on ... and easily traslated into ... (they are designed to be easily implemented in classic external memory models)

Having a ... eliminates .... (random accesses), access ordering

Not only are ..., but

The fingerprints are calculated on the fly during the scan of x ...

become so good that they are competitive with ...., and in some cases, even outperform them ...

concurrent programming is needed to make sure ....

designed to adapt to an evolving environment

most importantly

exchanges data between the computer presenting the applet and the computer serving it.

carries out sophisticated calculations

Presumably because

make ... capable of ...

greatly improve ... , it was still rather limited, though.

additional performance improvements

close the chapter with ...

gain

which pose ...

integrate with

In this chapter, we will ... how to ... and how to ...

we prefer the comfort of an integrated development environment, ( prefer the low cost of fingerprinting techinques)

space-hungry

begin exploring ... by ...

commnuicate the monmentous advances

extremely simple

check each number is present in the suffix array.

go into much more greater details

stand-alone algorithms for computing suffix arrays only and has been reused to

invoke/call

the mechanics of ...

If you have done sth, then when you do sth, you end up with ...

depend on the machine on which you will be running the java code ...

alleviate

The second method is invented to overcome the drawback of

in place of

as to/about

as far as

assign to a variable

stated goals

intermediate steps

tied to the behavior of

an assortment of

following the footsteps of

characters in the input string are processed from left to right

in greater detail

in the  order received

build up a whole new thing from scratch

In the interest of completeness, we briefly discuss the date and time formatting options of the printf method

immediately follow

preceding statement

tuned version

hang on to the object

track down

invisible to the remainder of the program

put it another way

>>>>

For full-text indexing of massive data, the suffix and LCP (longest common prefix) arrays have been recognized as fundamental data structures, and there are at least two needs in practice for checking their correctness, i.e. program debugging and verifying the arrays constructed by probabilistic algorithms. In this paper, we propose two probabilistic methods to check the suffix and LCP arrays of constant or integer alphabets in external memory by using a Karp-Rabin fingerprinting technique, where the checking result is wrong only with a negligible error probability. The first method checks the lexicographical order and the LCP-value of two suffixes by computing and comparing the fingerprints of their LCPs. This method is rather general in terms of that it can verify any full or sparse suffix/LCP array of any order. The second method is more space efficient, it first employs the fingerprinting technique to verify a subset of the given suffix and LCP arrays, from which then a copy of the suffix and LCP arrays is produced by using the induced sorting method and compared with the given arrays for verification, where the copy of the induced suffix and LCP arrays can be removed for constant alphabets.

@@

Our contribution comprises two methods to probabilistically verify any given suffix and LCP arrays. In principle, Method A checks the lexical order and the LCP-value of two neighboring suffixes in the given SA by literally comparing the characters of their LCPs. For reducing the time complexity of a comparison between two sequences of characters, we use a Karp-Rabin fingerprinting technique to convert each sequence into a single integer, called fingerprint, and compare the fingerprints instead to check the equality of two sequences. The algorithm for Method A involves multiple scans and sorts on sets of $\mathcal{O}(n)$ fixed-size items. Its implementation in external memory suffers from a space bottleneck due to the large disk volume taken by each sort. 
To overcome this drawback, Method B first employs the fingerprinting technique to check a subset selected from the given suffix and LCP arrays, then it utilizes the IS method to induce the final suffix and LCP arrays from the verified subset and literally compares them with the input arrays to ensure the correctness of the latter. Our experimental results indicate that the program for Algorithm~\ref{alg:2} designed by Method B only takes around half as much disk space as the program for Algorithm~\ref{alg:1} designed by Method A.

The remainder of this paper is organized as follows. Sections~\ref{sec:method1} and~\ref{sec:method2} describe the two methods and their algorithmic designs. Section~\ref{sec:experiment} conducts an experimental study for performance evaluation of our programs for these algorithms. Finally, Section~\ref{sec:conclusion} gives some concluding remarks.

@@

Assume that the alphabet $\Sigma$ is of a constant size, we can check $sa$ and $lcp$ without producing a copy of the final suffix and LCP arrays in Algorithm~\ref{alg:2}. The idea is to compare the induced suffix/LCP items with their corresponding items in $sa/lcp$ during the inducing process. Specifically, when a suffix/LCP item $v_1$ is induced into a bucket, we check if it is equal to the corresponding item $v_2$ in $sa/lcp$. If $v_1 = v_2$, then $v_2$ is correct and we further use this value to induce the remaining suffix/LCP items. We describe more details of the adapted inducing process as below, where $lp_1/lp_2$ and $sp_1/sp_2$ point to the next items to be visited in the L-type and S-type sub-buckets, respectively.

@@

We select these solutions because they are the current fastest available to us. A runtime breakdown of these solutions on the prefixes of "enwiki" is given in Table~\ref{tbl:3}, from which we see that Solution 2 is about two times faster than Solution 1 and twice as fast as ProgA. This phenomenon is mainly due to the high speed of pSAscan in the experiment. However, it is worthy of pointing out that both pSAscan and Sparse-$\phi$ are of the time and I/O complexities proportional to $n^2/M$. This is much higher than eSAIS and our checking algorithms when $n$ increases, and thus poses a strict limitation to the scalability of Solution 2. As reported in~\cite{Karkkainen2015}, when $n$ is considerably greater than $M$, eSAIS is much more time and I/O efficient than pSAscan. In this experiment, pSAscan builds the SA for "enwiki\_8g" in time double as that for "enwiki\_1g". For big $n$, it is more reasonable to compare the results of our programs with that of Solution 1.

@@

It should be noticed that our programs are coded for experimental study only. From engineering aspects, there is still a big margin for better implementation. For example, several algorithms for induced sorting a suffix and/or LCP arrays were proposed these years \cite{Nong15,Nong14,Bingmann12}, with different methods for solving the key problem of retrieving the preceding character of a sorted suffix in the inducing process. A recent work \cite{Karkkainen2017} for engineering these IS methods with some implementation optimizing techniques achieves a significant improvement over the previous results. As reported, the peak disk use is $7n$ for 32-bit integers. Because the inducing process is the performance bottleneck for Method B, it is reasonable to expect that a better engineering implementation of the method will yield a remarkable performance improvement. An optimized engineering of our methods is out of the scope of this paper and will be addressed elsewhere.

@@

The work presented in~\cite{Fischer11} describes how to compute the LCP array during S2-S3 of the induction phase. Specifically, for any two suffixes placed at the neighboring positions in $sa$, their LCP-value can be computed according to one of the following two cases in respect to whether or not they are inserted into the same sub-bucket: if yes, then the LCP-value is one greater than that of the two suffixes from which inducing them; otherwise, the LCP-value is zero. The key operation herein is to compute the LCP-value of the inducing suffixes at the same time when inserting the induced suffixes into $sa$. For example, in lines 10-21 of Fig.~\ref{fig:example2}, we induce ${\sf suf}(12)$ and ${\sf suf}(6)$ into the neighboring positions of ${\sf sa\_bkt_L(2)}$, respectively. If we keep recording the minimal over $lcp(0, 5]$, then we can obtain the LCP-value of the inducing suffixes ${\sf suf}(13)$ and ${\sf suf}(7)$ immediately after putting ${\sf suf}(6)$ into $sa$. This problem is modeled as a range minimum query in~\cite{Fischer11} and can be answered within amortized $\mathcal{O}(1)$ time.

@@
